Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2017-11-09 LaTeX/Turnover.tex
working out neo W invasion from equilA

For tight linkage between the ancestral sex-determining locus and the selected locus we can calculate each of these terms exactly, while for weak selection we take a Taylor series of the leading eigenvalue. With weak selection, the leading eigenvalue, $\lambda$, for any $k$, is given up to order $\epsilon^2$ by equation \eqref{eq:lambda_ESD_k}.

@@

\begin{subequations}\label{Binvasion}
\begin{align}
\lambda_{mA}&= \left[w_{a}^\male (1-\alpha^\male_\Delta) \right]^{-1}
\frac{w_{A}^\female }{w_{A}^\female}
\frac{\left[ w_{A}^\male (1+\alpha^\male_\Delta) w_{AA}^\female + 
w_{a}^\male (1-\alpha^\male_\Delta) w_{Aa}^\female (1+\alpha^\female_\Delta) \right]}
{ 2w_{AA}^\female } \\
\lambda_{ma}&= \left[w_{a}^\male (1-\alpha^\male_\Delta) \right]^{-1}
\frac{w_{a}^\female}{w_{A}^\female}
\frac{\left[  w_{A}^\male (1+\alpha^\male_\Delta) w_{Aa}^\female (1-\alpha^\female_\Delta)+
w_{a}^\male (1-\alpha^\male_\Delta) w_{aa}^\female \right]}
{ 2w_{AA}^\female } \\
\chi_{mA}&= \frac{1}{2} \left[w_{a}^\male (1-\alpha^\male_\Delta) \right]^{-1}
\frac{w_{A}^\female }{w_{A}^\female}
 \frac{\left[  w_{a}^\male (1-\alpha^\male_\Delta) w_{Aa}^\female (1 + \alpha^\female_\Delta) \right]}
{ w_{AA}^\female } \frac{R}{2}\\
\chi_{ma}&= \frac{1}{2} \left[w_{a}^\male (1-\alpha^\male_\Delta) \right]^{-1}
\frac{w_{a}^\female}{w_{A}^\female}
 \frac{\left[  w_{A}^\male (1+\alpha^\male_\Delta) w_{Aa}^\female (1-\alpha^\female_\Delta) \right]}
{  w_{AA}^\female} \frac{R}{2}
\end{align}
\end{subequations}

@@

Haploid selection impacts the spread of neo-W haplotypes in three ways.
Firstly, the zygotic sex ratio becomes male biased, $\zeta>1/2$, when the $a$ allele (which is fixed on the Y) is favoured during competition among male gametes or by meiotic drive in males.
Specifically, at equilibrium $(B)$, the sex ratio is $\zeta = w_a^\male (1-\alpha^\male_\Delta) / (2\bar{w}_{H}^\male)$ where $2\bar{w}_{H}^\male= \big[w_a^\male (1-\alpha^\male_\Delta) +  w_A^\male (1+\alpha^\male_\Delta) \big]$ has been canceled out in equations \eqref{Binvasion} to leave the term $\left[w_{a}^\male (1-\alpha^\male_\Delta) \right]^{-1}$. 
Male biased sex ratios facilitate the spread of a neo-W because neo-W alleles cause the zygotes that carry them to develop as the rarer, female, sex.

@@

\begin{subequations}\label{Ainvasion}
\begin{align}
\lambda_{mA}&= \frac{a}{b} \left[w_{AA}^\female w_{Aa}^\male w_A^\male (1+\alpha^\male_\Delta) \phi +  w_{Aa}^\female (1+\alpha^\female_\Delta) w_a^\male \frac{c}{d}\right] / (2 w_a^\female) \\
\lambda_{ma}&= \frac{a}{b} \left[ w_{Aa}^\female (1-\alpha^\female_\Delta) w_{Aa}^\male w_A^\male (1+\alpha^\male_\Delta) \phi +  w_{aa}^\female w_a^\male \frac{c}{d}\right] / (2 w_A^\female) \\
\chi_{mA}&= \frac{a}{b} \frac{R}{2} \left[   w_{Aa}^\female (1+\alpha^\female_\Delta) w_a^\male \frac{c}{d} \right] / w_a^\female\\ 
\chi_{ma}&= \frac{a}{b} \frac{R}{2} \left[ w_{Aa}^\female (1-\alpha^\female_\Delta) w_{Aa}^\male w_A^\male (1+\alpha^\male_\Delta) \phi \right] / w_A^\female
\end{align}
\end{subequations}

@@

\begin{subequations}
\begin{align}
a &= w_a^\female \phi + w_A^\female \psi \\
b &= w_{AA}^\female \phi \left[ w_{Aa}^\male w_A^\male (1+\alpha^\male_\Delta) \phi \right]+ w_{Aa}^\female \psi \left[ w_{Aa}^\male w_A^\male (1+\alpha^\male_\Delta) \phi + w_{AA}^\male w_a^\male \psi \right] + w_{aa}^\female \psi \left( w_{AA}^\male w_a^\male \psi \right)\\
%c &= w_{Aa}^\male (1-\alpha^\male_\Delta) \phi \left[ w_{Aa}^\male (1+\alpha^\male_\Delta) \phi \right] + w_{Aa}^\male\phi \left[ 2w_{AA}^\male + w_{aa}^\male (1+\alpha^\male_\Delta) \right] \psi + w_{aa}^\male \psi \left(2w_{AA}^\male \psi \right) \\
c &= 
w_{Aa}^\male (1+\alpha^\male_\Delta) \phi \left[ w_{Aa}^\male (1-\alpha^\male_\Delta) \phi + w_{aa}^\male \psi \right] +
2 w_{AA}^\male\psi \left[ w_{Aa}^\male \phi + w_{aa}^\male \psi \right]
 \\
d &= w_{Aa}^\male (1+\alpha^\male_\Delta) \phi + w_{aa}^\male \psi
\end{align}
\end{subequations}

>>>>

The general conditions for the invasion of a neo-sex-determining allele are given in the main text, in terms of the growth rates of the mutant haplotypes in the absence of recombination ($\lambda_{mi}$) and the rate that recombination destroys them ($\chi_{mi}$).
For tight linkage between the ancestral sex-determining locus and the selected locus we can calculate these terms explicitly (see below).
For weak selection we can take a Taylor series of the leading eigenvalue. 
The leading eigenvalue, $\lambda$, for any $k$, is given up to order $\epsilon^2$ by equation \eqref{eq:lambda_ESD_k}.

@@

\begin{subequations}\label{Binvasion}
\begin{align}
\lambda_{mA}&= \left[w_{A}^\male (1+\alpha^\male_\Delta) \right]^{-1}
\frac{w_{A}^\female }{w_{A}^\female}
\frac{\left[ w_{A}^\male (1+\alpha^\male_\Delta) w_{AA}^\female + 
w_{a}^\male (1-\alpha^\male_\Delta) w_{Aa}^\female (1+\alpha^\female_\Delta) \right]}
{ 2w_{AA}^\female } \\
\lambda_{ma}&= \left[w_{A}^\male (1+\alpha^\male_\Delta) \right]^{-1}
\frac{w_{a}^\female}{w_{A}^\female}
\frac{\left[  w_{A}^\male (1+\alpha^\male_\Delta) w_{Aa}^\female (1-\alpha^\female_\Delta)+
w_{a}^\male (1-\alpha^\male_\Delta) w_{aa}^\female \right]}
{ 2w_{AA}^\female } \\
\chi_{mA}&= \frac{1}{2} \left[w_{A}^\male (1+\alpha^\male_\Delta) \right]^{-1}
\frac{w_{A}^\female }{w_{A}^\female}
 \frac{\left[  w_{a}^\male (1-\alpha^\male_\Delta) w_{Aa}^\female (1 + \alpha^\female_\Delta) \right]}
{ w_{AA}^\female } \frac{R}{2}\\
\chi_{ma}&= \frac{1}{2} \left[w_{A}^\male (1+\alpha^\male_\Delta) \right]^{-1}
\frac{w_{a}^\female}{w_{A}^\female}
 \frac{\left[  w_{A}^\male (1+\alpha^\male_\Delta) w_{Aa}^\female (1-\alpha^\female_\Delta) \right]}
{  w_{AA}^\female} \frac{R}{2}
\end{align}
\end{subequations}

@@

Haploid selection impacts the spread of neo-W haplotypes in three ways.
Firstly, the zygotic sex ratio becomes male biased, $\zeta>1/2$, when the $a$ allele (which is fixed on the Y) is favoured during competition among male gametes or by meiotic drive in males.
Specifically, at equilibrium $(B)$, the fraction female is $1 - \zeta = w_A^\male (1+\alpha^\male_\Delta) / (2\bar{w}_{H}^\male)$ where $2\bar{w}_{H}^\male= \big[w_a^\male (1-\alpha^\male_\Delta) +  w_A^\male (1+\alpha^\male_\Delta) \big]$ has been canceled out in equations \eqref{Binvasion} to leave the term $\left[w_{A}^\male (1+\alpha^\male_\Delta) \right]^{-1}$. 
Male biased sex ratios facilitate the spread of a neo-W because neo-W alleles cause the zygotes that carry them to develop as the rarer, female, sex.

@@

\begin{subequations}\label{Ainvasion}
\begin{align}
\lambda_{mA}&= \frac{a}{b} \left[w_{AA}^\female w_{Aa}^\male w_A^\male (1+\alpha^\male_\Delta) \phi +  w_{Aa}^\female (1+\alpha^\female_\Delta) w_a^\male c \right] / (2 w_a^\female) \\
\lambda_{ma}&= \frac{a}{b} \left[ w_{Aa}^\female (1-\alpha^\female_\Delta) w_{Aa}^\male w_A^\male (1+\alpha^\male_\Delta) \phi +  w_{aa}^\female w_a^\male c \right] / (2 w_A^\female) \\
\chi_{mA}&= \frac{a}{b} \frac{R}{2} \left[   w_{Aa}^\female (1+\alpha^\female_\Delta) w_a^\male c \right] / w_a^\female\\ 
\chi_{ma}&= \frac{a}{b} \frac{R}{2} \left[ w_{Aa}^\female (1-\alpha^\female_\Delta) w_{Aa}^\male w_A^\male (1+\alpha^\male_\Delta) \phi \right] / w_A^\female
\end{align}
\end{subequations}

@@

\begin{subequations}
\begin{align}
a &= w_a^\female \phi + w_A^\female \psi \\
b &= w_{AA}^\female \phi \left[ w_{Aa}^\male w_A^\male (1+\alpha^\male_\Delta) \phi \right] + w_{Aa}^\female \psi \left[ w_{Aa}^\male w_A^\male (1+\alpha^\male_\Delta) + w_{aa}^\male w_a^\male \right] \phi + w_{aa}^\female \psi \left( w_{aa}^\male w_a^\male \psi \right)\\
%c &= w_{Aa}^\male (1-\alpha^\male_\Delta) \phi \left[ w_{Aa}^\male (1+\alpha^\male_\Delta) \phi \right] + w_{Aa}^\male\phi \left[ 2w_{AA}^\male + w_{aa}^\male (1+\alpha^\male_\Delta) \right] \psi + w_{aa}^\male \psi \left(2w_{AA}^\male \psi \right) \\
c &= w_{Aa}^\male (1-\alpha^\male_\Delta) \phi + 2w_{aa}^\male \psi
\end{align}
\end{subequations}
